---
title: "Covid19 vaccine study Fall2021"
author: "Nathan Abromovich"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males <- filter(Covid91,sex=="Man")
```

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```
  
The box plot above in males that where sick the drug worked for them.  The fact that we are seeing males that took the placebo medication got sick verses the healthy guys that were 50/50 split shows that they drug had a positive effect in the data given. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
  
The numerical data show that 66.9% of people that took the placebo got sick while only 33% of people that took the drug got sick.  This shows that there is a 2:1 spread and show the drug to be effective. 

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```
  
According to the Chi-Squared test, the p-value is very low at (2.2e-16).  This number indicates that we need to reject the null hypothesis and that there is a difference between the males that took the placebo and those whom took the drug.

## Calculating Effectiveness For Males


The effectiveness of Covid91 vaccine for males is 67.2%. The confidence interval for males is 63.1 to 70.9.

```{r}
Maleestimate <- Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <-Malesmod$conf.int[2]
Malelowprob <- Malelow / (1+Malelow)
Maleloweffect <- Malelowprob *100

Malehighprob <- Malehigh / (1+Malehigh)
Malehigheffect <- Malehighprob *100

Maleprod <- Maleestimate / (1+Maleestimate)
Maleeffect <- Maleprod * 100

options(digit=1)
```


# Females
```{r}
Woman <- filter(Covid91, sex=="Woman")
```

```{r}
barchartGC(~infected + treatment,data=Woman)
barchartGC(~infected + treatment,data=Woman, type="percent")

```
  
The box plot is showing that the drug isn't as effective in woman than in males.  It presents that approximately 58% of woman that took the drug still got sick verses about 42% of placebo users getting sick.  

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Woman)
rowPerc(table1)
colPerc(table1)
```
  
  The numerical summaries confirmed what we saw in the box plot.  58.1% of women that took the Covid91 drug still got sick in comparison to the placebo drug taking women on only got sick 41.8% of the time. 
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Womansmod <- fisher.test(table1)
```
According to the Chi-Squared test, the p-value for women is .0006. This number indicates that we need to reject the null hypothesis and that there is a difference between the women that took the placebo and those whom took the drug.  The drug is not effective compared tot he placebo.


## Calculating Effectiveness For Women
  
The Effectiveness of the Covid91 vaccine in women is 40.4% with a confidence level of 35.1 to 45.9.   

```{r}
Womanestimate <- Womansmod$estimate
Womanlow <- Womansmod$conf.int[1]
Womanhigh <-Womansmod$conf.int[2]
Womanlowprob <- Womanlow / (1+Womanlow)
Womanloweffect <- Womanlowprob *100

Womanhighprob <- Womanhigh / (1+Womanhigh)
Womanhigheffect <- Womanhighprob *100

Womanprod <- Womanestimate / (1+Womanestimate)
Womanseffect <- Womanprod * 100

options(digit=1)
```

# Drug Users
```{r}
Users <- filter(Covid91, DrugUser=="yes")
```

```{r}
barchartGC(~infected + treatment,data=Users)
barchartGC(~infected + treatment,data=Users, type="percent")

```
  
The box plot for the group studied called drug users show that the Covid91 vaccine is working.  Within the drug users that got the vaccine close to half of them got sick when compared to the drug users that got the placebo vaccine.     


### Numerical Summaries 

```{r}
table3 <- xtabs(~infected + treatment, data=Users)
rowPerc(table3)
colPerc(table3)
```
  
The numerical summary on the drug users group shows again that the Covid91 vaccine is work for them.  The numbers show that 33.3% of drug users that got the Covid91 vaccine got sick while 66.7% of the drug users who only got the placebo vaccine got sick.  This shows that within the drug user group that they are two times more likely to get sick if they didn't get the Covid91 vaccine. 


```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
Usersmod <- fisher.test(table3)
```
The p-value of the part of the study is 0.002273.  This number indicates that we need to reject the null hypothesis and that there is a difference between the drug users that took the placebo and those whom took the drug.  The drug in this group of drug users shows to be effective. 


 



```{r}
Usersestimate <- Usersmod$estimate
Userslow <- Usersmod$conf.int[1]
Usershigh <-Usersmod$conf.int[2]
Userslowprob <- Userslow / (1+Userslow)
Usersloweffect <- Userslowprob *100

Usershighprob <- Usershigh / (1+Usershigh)
Usershigheffect <- Usershighprob *100

Usersprod <- Usersestimate / (1+Usersestimate)
Userseffect <- Usersprod * 100

options(digit=1)
```
## Calculating Effectiveness For Drug Users
  
The effectiveness of the Covid91 vaccine among drug users is 64.7% with a confidence level of 54 to 73.6.


# LGBTQ
```{r}
Gays <- filter(Covid91, LGBTQ=="gay")
```

```{r}
barchartGC(~infected + treatment,data=Gays)
barchartGC(~infected + treatment,data=Gays, type="percent")

```
  
The data in the above box plot shows that the Covid91 vaccine does not help in the LGBTQ group of people.  This is shown due to the amount of people that took the drug and still got sick while less than half of the people in this group took the placebo got sick when compared to the vaccine group.

## Numerical Summeries 

```{r}
table3 <- xtabs(~infected + treatment, data=Gays)
rowPerc(table3)
colPerc(table3)
```
  
The numerical summery of the gays group backs up the data we see in the box plot run earlier.  71% of the gays that took the Covid91 vaccine still ended up getting the virus while the 30% of gays that took the placebo got sick.  It shows that the COVID91 vaccine does not work in the LGBTQ group. 

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
Gaysmod <- fisher.test(table3)
```
The gays data shows a p-value of (5.741e-06).  This would tell us that we will reject the null hypothesis and that there is a difference between the LGBTQ group that took the COVID91 vaccine and the part of the LGBTQ group that took the placebo drug.  The COVID 91 vaccine is ineffective in the group of LGBTQ people.  


## Calculating Effectiveness For LGBTQ Group
  
The Covid91 vaccine effectiveness in the gays group is very low at only 28.6% effective with a confidence level of 20.5 to 38. 


```{r}
Gaysestimate <- Gaysmod$estimate
Gayslow <- Gaysmod$conf.int[1]
Gayshigh <-Gaysmod$conf.int[2]
Gayslowprob <- Gayslow / (1+Gayslow)
Gaysloweffect <- Gayslowprob *100

Gayshighprob <- Gayshigh / (1+Gayshigh)
Gayshigheffect <- Gayshighprob *100

Gaysprod <- Gaysestimate / (1+Gaysestimate)
Gayseffect <- Gaysprod * 100

options(digit=1)
```

# Conclusions and recommendations

In conclusion we tested four groups of people.  The groups being males, females, drug users and people that are LGBTQ.  In each group we looked at created box plots to see how the COVID91 treatment compared to placebo in people that got sick and stayed healthy.  Numerical summeries, Chi-squared test and Fisher exact test helped us break that data down even further to be able to compute the COVID91 drug effectiveness for each group tested.

A strong recommendation only based on the data received would be to only take the COVID91 vaccine if you are in the LGBTQ group.  The LGBTQ group tested is the only group that the COVID91 vaccine was effective in.  All other groups, males, females and drug users showed an ineffectiveness to the COVID91 vaccine with the drug users being the least effective group. 


#  **Disclaimer** 

This study does not provide medical advise.  Please see your doctor if you think you have COVID91.



